
```{r}
#| label: p-val-test-cgg
permuted_test_stats = collect_permutation_test_stats(
  output_folder = output_folder,
  permuting_variables = "FX3")

observed_test_stat = get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c('under100', 'over100'))

pval_cgg_gp34_v1 = permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats,
  n_permutations = n_permutations
)

test_cgg_summary <- pval_cgg_gp34_v1 |> 
  summary(N = nrow(v1_usable_cases))
```

```{r}
#| eval: false
#| label: plot-permutation-test-cgg
plot_permutation_results(
  permuted_test_stats,
  observed_test_stat)
```


When we compared
between 55-99 vs 100-200 CGG repeats in males and females combined,
no significant difference was found
(p = `r pval_cgg_gp34_v1`)
(@suppfig-pvd-by-cgg).
The finding of only minimal differences
between 55-99 vs 100-200 CGG repeats is surprising,
since several studies have found that the higher the CGG repeats,
the earlier the onset and the faster the progression of FXTAS
[@greco2006neuropathology; @leehey2008fmr1; @tassone2023insight].
Perhaps the cut-off of 100 is too high for this distinction to be made,
and thus searching for an appropriate cut-off of CGG repeats is warranted 
to determine if a specific threshold of CGG repeat size 
triggers the progression of FXTAS.
