::: {#suppfig-pvd_chosen_subtypes_bumpplot}

```{r}
#| fig-height: 9
#| fig-width: 15
color_info <- tibble::tribble(
  ~event_name, ~line_color, ~highlight_color,
  "MRI: Cerebellar: Mild",  "orange", "#FFF",
  "MRI: Cerebral: Moderate/Severe", "orange", "#FFF",
  # "SCID: Anxiety Disorders: Threshold", "orange", "#FFF",
  "SCID: Mood Disorders: Threshold", "orange", "#FFF",
  "Corpus Callosum-Thickness: Thin", "orange", "#FFF",
  "Splenium WM Hyperintensity: Mild", "orange", "#FFF",
  "Genu WM Hyperintensity", "orange", "#FFF",
  "Parkinsonian Features", "green", "#FFF",
  "Parkinsons", "green", "#FFF",
  "FXTAS Stage: 1", "black", "yellow",
  "FXTAS Stage: 2", "black", "yellow",
  "FXTAS Stage: 3", "black", "yellow",
  "FXTAS Stage: 4", "black", "yellow",
  "FXTAS Stage: 5", "black", "yellow",
  "SWM Between Errors: > 26", "purple", "#FFF",
  "Postural Tremor", "red", "#FFF",
  "Intermittent Tremor", "brown", "#FFF",
  "Intention Tremor", "brown", "#FFF",
  "Head Tremor", "cyan3", "#FFF",
  "Tandem Walk: Abnormal (<10 Steps)", "magenta", "#FFF",
  "Ataxia Severity: 1", "magenta", "#FFF",
  
)

figs |> pvd_subtype_lineplot(
  smooth = 30,
  y_lab = "",
  align_stage = FALSE,
  events_to_highlight = color_info)
```


**Differences in event sequences between latent subtypes.**
Lines are used to show patterns of clinical interest.
Text colors indicate symptoms grouped by measurement type.

:::

{{< pagebreak >}}

:::{#suppfig-bump-aligned}

```{r}
#| fig-height: 9
#| fig-width: 15
figs |> pvd_subtype_lineplot(
  smooth = 30,
  y_lab = "",
  align_stage = TRUE,
  events_to_highlight = color_info)
```

**Differences in event sequences between latent subtypes,**
**aligned by FXTAS stage.**
Lines are used to show patterns of clinical interest.
Text colors indicate symptoms grouped by measurement type.

:::

