
```{r}
#| label: p-val-test-cgg-male

permuted_test_stats <- collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "Male",
  permuting_variables = "FX3"
)

observed_test_stat <- get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c("under100_Male", "over100_Male")
)

pval_male_cgg_gp34_v1 <- permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats,
  n_permutations = n_permutations
)

test_male_cgg_summary <- pval_male_cgg_gp34_v1 |>
  summary(N = nrow(v1_usable_male_cases))
```

```{r}
#| eval: false
#| label: plot-permutation-test-cgg-male
plot_permutation_results(
  permuted_test_stats,
  observed_test_stat)
```

```{r}
#| label: p-val-test-cgg-female

permuted_test_stats <- collect_permutation_test_stats(
  output_folder = output_folder,
  stratifying_levels = "Female",
  permuting_variables = "FX3"
)

observed_test_stat <- get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c("under100_Female", "over100_Female")
)

pval_female_cgg_gp34_v1 <- permutation_test(
  observed_test_stat = observed_test_stat,
  permuted_test_stats = permuted_test_stats,
  n_permutations = n_permutations
)

n_over_100_female <-
  v1_usable |>
  dplyr::filter(
    Gender == "Female",
    CGG >= 100
  ) |>
  nrow()

N_female <- v1_usable_cases |> filter(Gender == "Female") |> nrow()
test_female_cgg_summary <- pval_female_cgg_gp34_v1 |>
  summary(N = nrow(v1_usable_female_cases))
```

```{r}
#| eval: false
#| label: plot-permutation-test-cgg-female
plot_permutation_results(
  permuted_test_stats,
  observed_test_stat)
```

We did not find significant evidence of an overall difference between CGG <100
and CGG ≥100 among males (p = `r pval_male_cgg_gp34_v1`)
([Supplementary Fig. @suppfig-pvd-by-cgg-male-1]).
SCID anxiety disorders and SWM between errors occurred
prior to Stage 1 (1st and 2nd in sequence) in males with CGG repeats <100,
whereas they occurred between stage 4 and 5 (32nd and 27th in sequence)
in males with CGG repeats ≥ 100
([Supplementary Fig. @suppfig-pvd-by-cgg-male-2]).
Postural,
resting,
and
intermittent tremor
occurred in earlier FXTAS stages in those with CGG repeats <100
compared to those with CGG repeats ≥100.
SCID substance use disorders and somatoform disorders
and MRI biomarkers
occurred later in the event sequence in participants with CGG repeats <100
compared to those with CGG repeats ≥100.
We also did not find a significant difference between CGG <100 and CGG ≥100
among females (p = `r pval_female_cgg_gp34_v1`)
(@suppfig-pvd-by-cgg-female).
It should be noted that this comparison lacked statistical power
given the limited sample size of only
`r n_over_100_female` females with CGG ≥100.
