
::: {#suppfig-pvd-by-cgg-male-1 custom-style="Caption"}

```{r}
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width
#| label: pvd-males-by-cgg

# compact pvd
list(
    "Males, CGG &lt; 100:" = fig_under100_males,
    "Males, CGG &ge; 100:" = fig_over100_males
  ) |>
plot_compact_pvd(
  scale_colors = scale_colors1, 
  y_text_size = y_text_size
)

```

**Event sequences stratified by CGG repeats, among males.**
`r compact_fig_caption`
`r test_male_cgg_summary`
:::

{{< pagebreak >}}

:::{#suppfig-pvd-by-cgg-male-2 custom-style="Caption"}

```{r}
#| fig-height: !expr pvd_line_height
#| fig-width: !expr pvd_line_width
#| label: lineplot-by-cgg-male
list(
    "Males, CGG &lt; 100" = fig_under100_males,
    "Males, CGG &ge; 100" = fig_over100_males
  ) |>
  pvd_lineplot(text_size = lineplot_text_size)
```

**Positional differences in estimated event sequence**
**between CGG repeat levels, among males**.
`r compact_fig_cap2`
`r test_male_cgg_summary`
:::


{{< pagebreak >}}

::: {#suppfig-pvd-by-cgg-female-1 custom-style="Caption"}

```{r}
#| label: supp-fig-pvd-by-cgg-female-1
#| fig-height: !expr pvd_height
#| fig-width: !expr pvd_width

# compact pvd
list(
    "Females, CGG &lt; 100:" = fig_under100_females,
    "Females, CGG &ge; 100:" = fig_over100_females
  ) |> 
  plot_compact_pvd(scale_colors = scale_colors1, y_text_size = y_text_size)

```


**Event sequences stratified by CGG repeats (<100 vs 100+), among females.** 
`r compact_fig_caption`
`r test_female_cgg_summary`
:::

{{< pagebreak >}}

:::{#suppfig-pvd-by-cgg-female-2 custom-style="Caption"}

```{r}
#| label: supp-fig-pvd-by-cgg-female-2
#| fig-height: !expr pvd_line_height
#| fig-width: !expr pvd_line_width

# lineplot of event sequences

list(
  "Females, CGG &lt; 100" = fig_under100_females,
  "Females, CGG &ge; 100" = fig_over100_females
) |> 
  pvd_lineplot(text_size = lineplot_text_size)
  
```

**Positional differences in estimated event sequence**
**between CGG repeats (<100 vs 100+)**. 
`r compact_fig_cap2`
`r test_female_cgg_summary`
:::
