
```{r}
#| label: p-val-test-sex

permuted_test_stats_sex <- collect_permutation_test_stats(
  output_folder = output_folder,
  permuting_variables = "Gender"
)

observed_test_stat_sex <- get_observed_permutation_test_stat(
  output_folder,
  dataset_names = c("females", "males")
)

pval_sex_gp34_v1 <- permutation_test(
  observed_test_stat = observed_test_stat_sex,
  permuted_test_stats = permuted_test_stats_sex,
  n_permutations = n_permutations
)

pval_sex_gp34_v1_formatted <- pval_sex_gp34_v1 |> scales::label_pvalue()()

test_sex_summary <- pval_sex_gp34_v1 |> 
  summary(N = nrow(v1_usable_cases))

```


```{r}
#| eval: false
#| label: plot-permutation-test-sex
plot_permutation_results(
  permuted_test_stats,
  observed_test_stat)
```


We first characterized 
sequential orders of FXTAS symptoms by sex 
and sex differences in the estimated sequential orders.
@fig-first-only-pvd shows 
the estimated sequential orders
of FXTAS `r outcomes_name`s 
in males and females, 
and @fig-first-only-diffs highlights sex differences in those orders.
@suppfig-stage-barplot-by-sex shows the 
number of premutation carriers at each disease progression stage, 
stratified by sex.
We found statistically significant evidence of a 
difference in sequences of `r events_name`
between males and females (p `r pval_sex_gp34_v1_formatted`). 
It appears that 
female premutation carriers developed 
mood disorder symptoms prior to FXTAS Stage 1 diagnosis,
whereas male carriers experienced these symptoms much later 
(after FXTAS Stage 4) (@fig-first-only-diffs).
Postural tremors developed between Stage 2 and 3 for females,
and between Stages 3 and 4 for males.
Several MRI biomarkers of decline
also appear to begin at earlier FXTAS stages for males,
except for mild MRI cerebral biomarkers, which began earlier in females.
