:::{#suppfig-cvic-1 custom-style="Caption"}

```{r}
#| label: plot-cvic
sustain_output_CV$CVIC |> plot_CVIC()
```

**Selection criteria for number of latent subtypes:**
**cross-validation information criterion (CVIC)**.
Each data point represents the estimated CVIC (y-axis)
for the model fit with a given number of latent subtypes (x-axis)
on the full data set 
(N = `r nrow(v1_usable_cases)` carriers
and 
`r nrow(v1_usable_controls)` controls).

:::

:::{#suppfig-cvic-2 custom-style="Caption"}

```{r}
#| label: plot-ooffl
sustain_output_CV$loglike_matrix |> plot_cv_loglik(y_text_size = y_text_size)
```

**Selection criteria for number of latent subtypes:**
**Out-Of-Fold Log-Likelihood criterion (OOFLL)**.
Each data point represents the OOFLL statistic
(i.e., the test set log-likelihood across cross-validation folds)
for the Ordinal SuStaIn model fit with a given number of latent subtypes
on the full data set
(N = `r nrow(v1_usable_cases)` carriers
and 
`r nrow(v1_usable_controls)` controls).
:::
